\large How resource abundance and stochasticity affect organisms' space-use requirements

### 2024-03-28 writing/manuscript.Rmd
updated appendices and methods for empirical model

\noindent Fig. \@ref(fig:tapir-mw) illustrates how a tapir in the Brazilian Cerrado [data from the individual labelled as "Anna" in the dataset from @medici_movement_2022] adapts its spatial needs to changes in $\e(R)$ and $\var(R)$. Panels A and B show the changes in seven-day average mean and variance in NDVI, respectively, experienced by the tapir during the tracking period. The mean and variance in NDVI were estimated using a Generalized Additive Model for Location and Scale [GAMLS, theory: @rigby_generalized_2005; @stasinopoulos_generalized_2007, examples: @bjorndahl_abrupt_2022; @marien_timing_2022; @gushulak_impacts_2024] with a Beta family of distributions (NDVI values ranged from 0.3534 to 0.9475). Panel C shows the changes in the tapir's 7-day home range over time. Note how the tapir uses more space during periods of lower NDVI (e.g., August 2017) and less space during periods with high NDVI (January 2018). Additionally, when resources are scarce and highly unpredictable (August 2018), the tapir uses up to 5 times more space than when resources are abundant and predictable (e.g., January 2018). Finally, panels D and E show the estimated (marginal) effects of $\mu(t)$ and $\sigma^2(t)$ on the tapir's space use. Since $\mu(t)$ and $\sigma^2(t)$ are strongly correlated (panel F) and (spatio)temporally autocorrelated (panels A and B), the effects of $R$ on $H$ should be modeled carefully. To avoid over-fitting the model, we constrained the smooth effects of $\mu(t)$ and $\sigma^2(t)$ using a shape-constrained additive model [@pya_scam_2022] with monotone deceasing and increasing P-splines for $\mu(t)$ and $\sigma^2(t)$, respectively. Additional information is provided in appendix C. Overall, the results presented in fig. \@ref(fig:tapir-mw) support our hypotheses, as the taprir's 7-day home range decreases nonlinearly with $\mu(t)$ and increases nonlinearly with $\sigma^2(t)$. However, due to the strong nonlinear correlation between the two predictors, the marginal effects do not follow the data closely, particularly in the case of $\mu(t)$. We provide comments about these results in the related section in the discussion below.

>>>>

\noindent Fig. \@ref(fig:tapir-mw) illustrates how a tapir in the Brazilian Cerrado [data from the individual labelled as "Anna" in the dataset from @medici_movement_2022] adapts its spatial needs to changes in $\e(R)$ and $\var(R)$. Panels A and B show the changes in seven-day average mean and variance in NDVI, respectively, experienced by the tapir during the tracking period. The mean and variance in NDVI were estimated using a Generalized Additive Model for Location and Scale [GAMLS, theory: @rigby_generalized_2005; @stasinopoulos_generalized_2007; examples: @bjorndahl_abrupt_2022; @marien_timing_2022; @gushulak_impacts_2024] with a Beta family of distributions (NDVI values ranged from 0.3534 to 0.9475). Panel C shows the changes in the tapir's 7-day home range over time. Note how the tapir uses more space during periods of lower NDVI (e.g., August 2017) and less space during periods with high NDVI (January 2018). Additionally, when resources are scarce and highly unpredictable (August 2018), the tapir uses up to 5 times more space than when resources are abundant and predictable (e.g., January 2018). Finally, panels D and E show the estimated (marginal) effects of $\mu(t)$ and $\sigma^2(t)$ on the tapir's space use. Since $\mu(t)$ and $\sigma^2(t)$ are strongly correlated (panel F) and (spatio)temporally autocorrelated (panels A and B), the effects of $R$ on $H$ should be modeled carefully. To avoid over-fitting the model, we constrained the smooth effects of $\mu(t)$ and $\sigma^2(t)$ using low number of knots ($\tt{k = 3}$). Additional information is provided in appendix C. Overall, the results presented in fig. \@ref(fig:tapir-mw) support our hypotheses, as the taprir's 7-day home range decreases nonlinearly with $\mu(t)$ and increases nonlinearly with $\sigma^2(t)$. However, due to the strong nonlinear correlation between the two predictors, the marginal effects do not follow the data closely, particularly in the case of $\mu(t)$. We provide comments about these results in the related section in the discussion below.
